The Birth of a Galaxy: Population III Metal Enrichment and Population II Stellar Populations

### 2010-10-25 science_ms.tex
Added some discussion on the ionized bubble (estimated) radius around the z=8.6 galaxy and connection to DLAs.

In this Report, we focus on the birth of two galaxies prior to
reionization with a cosmological AMR radiation hydrodynamics
simulation.  Supernovae from Pop III stars provide the necessary heavy
elements for the transition to a Population II stellar population,
which we have directly simulated.  These two galaxies have a 10--15\%
probability in surviving as present-day ``fossil'' galaxies
\cite{Gnedin06}, otherwise they will be incorporated into galactic
stellar halos.

We find that one PISN is sufficient to enrich the star-forming halo
and surrounding $\sim 5$ kpc to a metallicity of 10$^{-3} Z_\odot$,
given $M_{\rm char} = 100~\Ms$.  If the first stars have a lower
characteristic mass that favor hypernovae \cite{Tumlinson07_IMF}, then
this metallicity floor should be lowered by a factor of $\sim 10$
because (1) the metal ejecta is lowered by a factor of $\sim 50$ and
(2) the mixing mass is approximately decreased by a factor of $(E_{\rm
  hyp}/E_{\rm PISN})^{3/5}$ in the Sedov-Taylor solution, where
$(E_{\rm hyp}/E_{\rm PISN}) \sim 0.1$ is the ratio of explosion
energies of a hypernova and PISN.  In the case where this metallicity
floor is less than the critical metallicity, then the next instance of
SF will further enrich the ISM, solidifying the transition to Pop II
SF.  We conclude that it only takes one, at most two, SNe from Pop III
stars in the halo progenitors to complete the transition to Population
II \cite{Frebel10}.  The question of whether the critical metallicity
is \tento{-6} or \tento{-3.5} $Z_\odot$ is most applicable to nearby
halos where the heavy elements mix slowly into dense cores as the
blastwave overtakes it.  Less than two percent of stars have [Z/H] $<
-3$ in both systems, consistent with observations of metal-poor stars
in the halo and dSphs \cite{Beers05, Battaglia10}.

>>>>

In this Report, we focus on the birth of two galaxies prior to
reionization with a cosmological AMR radiation hydrodynamics
simulation.  Supernovae from Pop III stars provide the necessary heavy
elements for the transition to a Population II stellar population,
which we have directly simulated.  These two galaxies have a 10--15\%
probability in surviving as present-day ``fossil'' galaxies
\cite{Gnedin06}, otherwise they will be incorporated into galactic
stellar halos.  A $z=8.55$ galaxy was recently spectroscopically
confirmed that is contained in an ionized bubble with radius 0.1--0.5
Mpc with an uncharacterized population of galaxies contributing to the
local ionizing radiation field \cite{Lehnert10_z8.6}.  The galaxies
simulated here might represent this population.

We find that one PISN is sufficient to enrich the star-forming halo
and surrounding $\sim 5$ kpc to a metallicity of 10$^{-3} Z_\odot$,
given $M_{\rm char} = 100~\Ms$.  DLA systems have a metallicity floor
on the same order \cite{DLAs}, and metal enrichment from Pop III SNe
provides a possible explaination.

If the first stars have a lower characteristic mass that favor
hypernovae \cite{Tumlinson07_IMF}, then this metallicity floor should
be lowered by a factor of $\sim 10$ because (1) the metal ejecta is
lowered by a factor of $\sim 50$ and (2) the mixing mass is
approximately decreased by a factor of $(E_{\rm hyp}/E_{\rm
  PISN})^{3/5}$ in the Sedov-Taylor solution, where $(E_{\rm
  hyp}/E_{\rm PISN}) \sim 0.1$ is the ratio of explosion energies of a
hypernova and PISN.  In the case where this metallicity floor is less
than the critical metallicity, then the next instance of SF will
further enrich the ISM, solidifying the transition to Pop II SF.  We
conclude that it only takes one, at most two, SNe from Pop III stars
in the halo progenitors to complete the transition to Population II
\cite{Frebel10}.  The question of whether the critical metallicity is
\tento{-6} or \tento{-3.5} $Z_\odot$ is most applicable to nearby
halos where the heavy elements mix slowly into dense cores as the
blastwave overtakes it.  Less than two percent of stars have [Z/H] $<
-3$ in both systems, consistent with observations of metal-poor stars
in the halo and dSphs \cite{Beers05, Battaglia10}.
